Adaptive Sequential Model Selection

### 2015-07-10 paper.tex
rob changes

By contrast, the LASSO path does not in general add the variable with the largest $t$-statistic. In that case, we could use instead the {\em selective next-$\lambda$ test}, which uses test statistic
\begin{equation}
  U_k = \inf\{ \lambda:\; \hat\beta^\lambda \notin \beta(E_{k-1})\}.
\end{equation}
Here $\beta(E_{k-1})$ is the support of the ever-active set $E_{k-1}$, so that $U_k$ is the next $\lambda$ at which a new predictor is added by the lasso.
\RTcomment{Will: we never define $\beta(E)$: have i said it right?}
The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}

@@

\subsection{An analytic approximation for the max-t test}
 Suppose we have taken $k$ steps of forward stepwise regression with $p$ predictors and let the selected model  
 be $E$. Suppose we choose predictor $j^*$ at the next step.

@@

Figure \ref{fig:approxpv} shows an example with $n=50$ and $p=10$ features having pairwise correlation 0.5.
There are two moderate signals.
Shown are the p-values from the ``exact'' numerical computation (with $100,000$ bootstrap samples)
and the approximation based on (\ref{eqn:approxpv}).
\begin{figure}[htp]
\centering
  \includegraphics[width=3in]{figs/approxpv.pdf}
  \caption{Exact max-t pvalues and approximation based on (\ref{eqn:approxpv}), for a simulated example
  with $n=50, p=10$.}
  \label{fig:approxpv}
\end{figure}
The approximation seems quite good. By Sidak's lemma, we might conjecture that the approximation is conservative, but we have no proof of this.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
  Method, Stopping rule& $p_{\text screen}$ & FWER$_{\text mod}$ & FDR$_{\text mod}$ &FDR$_{\text var}$ \\
% & \$p\_\{$\backslash$text\{screen\}\} & \$$\backslash$text\{FWER\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{variable\}\}\$ \\ 
  \hline
Nominal, Simple & 0.832 & 0.252 & 0.097 & 0.171 \\ 
  Nominal, Forward & 0.931 & 0.896 & 0.345 & 0.426 \\ 
  Nominal, Strong & 0.003 & 0.000 & 0.000 & 0.035 \\ 
  \hline
  MaxT, Simple & 0.369 & 0.000 & 0.002 & 0.042 \\ 
  MaxT, Forward & 0.668 & 0.038 & 0.026 & 0.080 \\ 
  MaxT, Strong & 0.025 & 0.006 & 0.004 & 0.038 \\ 
  \hline
  MaxT-Identify, Simple & 0.404 & 0.000 & 0.002 & 0.042 \\ 
  MaxT-Identify, Forward & 0.688 & 0.039 & 0.026 & 0.082 \\ 
  MaxT-Identify, Strong & 0.024 & 0.003 & 0.002 & 0.037 \\ 
  \hline
  Saturated, Simple & 0.000 & 0.000 & 0.000 & 0.029 \\ 
  Saturated, Forward & 0.013 & 0.001 & 0.000 & 0.030 \\ 
  Saturated, Strong & 0.000 & 0.000 & 0.000 & 0.033 \\ 
  \hline
  Knockoffs & 0.252 &  &  & 0.103 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

By contrast, the LASSO path does not in general add the variable with the largest $t$-statistic. In that case, we could use instead the {\em selective next-$\lambda$ test}, which uses test statistic
\begin{equation}
  U_k = \inf\{ \lambda:\; \hat\beta^\lambda \notin \beta(E_{k-1})\}.
\end{equation}
Here $\beta(E_{k-1})$ is the support of the ever-active set $E_{k-1}$, so that $U_k$ is the next $\lambda$ at which a new predictor is added by the lasso.
\RTcomment{Will: we never define $\beta(E)$: have i said it right?}
The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}
\begin{figure}
  % source code: bivariateSelVSat.R
  \includegraphics[width=\textwidth]{figs/larexample.pdf}
  \label{fig:larexample}
\end{figure}

@@

\begin{comment}
\subsection{An analytic approximation for the max-t test}
 Suppose we have taken $k$ steps of forward stepwise regression with $p$ predictors and let the selected model  
 be $E$. Suppose we choose predictor $j^*$ at the next step.

@@

Figure \ref{fig:approxpv} shows an example with $n=50$ and $p=10$ features having pairwise correlation 0.5.
There are two moderate signals.
Shown are the p-values from the ``exact'' numerical computation (with $100,000$ bootstrap samples)
and the approximation based on (\ref{eqn:approxpv}).
\begin{figure}[htp]
\centering
  \includegraphics[width=3in]{figs/approxpv.pdf}
  \caption{Exact max-t pvalues and approximation based on (\ref{eqn:approxpv}), for a simulated example
  with $n=50, p=10$.}
  \label{fig:approxpv}
\end{figure}
The approximation seems quite good. By Sidak's lemma, we might conjecture that the approximation is conservative, but we have no proof of this.
\end{comment}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
  Method, Stopping rule& $p_{\text screen}$ & FWER$_{\text mod}$ & FDR$_{\text mod}$ &FDR$_{\text var}$ \\
% & \$p\_\{$\backslash$text\{screen\}\} & \$$\backslash$text\{FWER\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{variable\}\}\$ \\ 
  \hline
Nominal, Simple & 0.832 & 0.252 & 0.097 & 0.171 \\ 
  Nominal, Forward & 0.931 & 0.896 & 0.345 & 0.426 \\ 
  Nominal, Strong & 0.003 & 0.000 & 0.000 & 0.035 \\ 
  \hline
  MaxT, Simple & 0.369 & 0.000 & 0.002 & 0.042 \\ 
  MaxT, Forward & 0.668 & 0.038 & 0.026 & 0.080 \\ 
  MaxT, Strong & 0.025 & 0.006 & 0.004 & 0.038 \\ 
  \hline
  MaxT-Identify, Simple & 0.404 & 0.000 & 0.002 & 0.042 \\ 
  MaxT-Identify, Forward & 0.688 & 0.039 & 0.026 & 0.082 \\ 
  MaxT-Identify, Strong & 0.024 & 0.003 & 0.002 & 0.037 \\ 
  \hline
  Saturated, Simple & 0.000 & 0.000 & 0.000 & 0.029 \\ 
  Saturated, Forward & 0.013 & 0.001 & 0.000 & 0.030 \\ 
  Saturated, Strong & 0.000 & 0.000 & 0.000 & 0.033 \\ 
  \hline
  Knockoffs & 0.252 &  &  & 0.103 \\ 
   \hline
\end{tabular}
\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
\label{tab:stopping}
\end{table}
